Theoretical prediction of disrupted Min oscillation in flattened \emph{Escherichia coli}

### 2015-09-09 paper/plos.tex
remove \changed uses from last round with referees.
respond to last referees

@@

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step. 
% Author Summary not valid for PLOS ONE submissions.   
\section*{Author Summary}
  When an E. coli cell divides it must split into two relatively equal
  volumes in order to survive.  One mechanism that aids in this
  process is a system of proteins that oscillate between cell ends and
  help the original cell divide along its center.  A recent experiment
  found that by drastically deforming these cells this oscillation can
  be disrupted.  We examine these oscillations in similarly deformed
  cells computationally, using two widely used simulation models, and
  find good agreement using one of them.  The other model fails in a
  manner far more dramatic than has been previously observed.
  Furthermore, we find that the effect observed experimentally in
  asymmetric and flattened cells is primarily a result of the
  flattening and lateral expansion, not the asymmetry.

@@

Several models \cite{bonny2013membrane,
  halatek2012highly} modify Huang's original model to more accurately
reflect experimental findings, which were published after his 2003
paper~\cite{meacci2006mobility, loose2011min}. These models introduce
diffusion along the membrane that is two to three orders of magnitude
slower than that within the cytoplasm, and increase the rate of
cytoplasmic diffusion by a factor of about five, reflecting an
experimental measurement of these
parameters~\cite{meacci2006mobility}.  They also increase
or assume instantaneous~\cite{bonny2013membrane} the nucleotide
exchange rate.  In addition, they revise the process of MinD
recruitment to the membrane, removing recruitment of MinD to the
membrane by MinD-MinE complexes, behavior that has not been observed
experimentally.  The model of Bonny \emph{et al.} further more allows
MinE to remain independently attached to the membrane after
disassociating with MinD~\cite{bonny2013membrane}.
%

@@

%% biology~\cite{lawson2013spatial, robb2014stochastic,
%%   oguz2014stochastic, fu2013deterministic, rudiger2014stochastic}, and
In this paper we use Huang's original 2003
model~\cite{huang2003dynamic}, which has been widely
used~\cite{hattne2005stochastic,huang2004min,kerr2006division,
  fange2006noise} in both deterministic and stochastic variants, to
study flattened cells of $0.4\micron$ thickness that are similar to
those observed by M\"annik \emph{et al.}~\cite{mannik2009bacterial}.
This allows us to test the performance of this important model in more
extreme situations than has been done previously, furthering our
understanding of how and to what extent it breaks down.

@@

However Fig.~\ref{randst-plot-ave} leaves unclear the importance of
the flattening and accompanying lateral expansion of the cell versus
cell shape irregularity and asymmetry in the temporal regularity of
the oscillations.  Although in the movies the stadiums appear to
display, on average over long time scales, a somewhat more regular
oscillation than appears in the irregular shapes \emph{A} and
\emph{B}, the limited time range of the arrow plots (350 seconds) is
inadequate to make this comparison.  We therefore return to the
correlation function between the number of proteins in a segment at
the top of the cell and the number of proteins in a segment at the
bottom of the cell.  The ``top'' and ``bottom'' regions for computing
the correlation time were established by dividing the long axis into
thirds.  This resulted in regions at the top and bottom ranging from
19\% to 33\% of the cell volume for shapes \emph{A} and \emph{B}, with
the stadium shapes featuring 30\% of the volume in the top and bottom
regions.  Fig.~\ref{fig:corr-B} shows this plot for shape \emph{B} and
stadium \emph{B}.  Here we plot correlations of correlation times up
to 500 seconds, a biologically relevant timescale.  E. coli cell
doubling times vary widely depending on environmental conditions, but
will often be between 20 minutes to 100
minutes~\cite{pierucci1972chromosome}, while fluorescent microscopy
experiments have shown that the FtsZ ring builds itself with a half
life of roughly 30 seconds~\cite{stricker2002rapid}.

@@

\section*{Conclusion}
We find qualitative agreement between our stochastic simulations and
experiments showing disrupted bipolar MinD oscillation in asymmetric
flattened E. coli cells~\cite{mannik2012robustness}.  As observed
experimentally, our simulations predict that MinD maxima will form in
a spatially irregular sequence of locations, which follow an overall
bipolar trend.  This result builds upon existing results showing that
this model and its variants are effective in a variety of wild-type
and mutated cell shapes~\cite{fange2006noise, varma2008min,
  kruse2007experimentalist}, and reinforces that stochastic variations
of Huang's 2003 model~\cite{fange2006noise, kerr2006division} has
considerable predictive power beyond the standard wild-type cell.

>>>>

\newcommand{\changed}[1]{{\color{red}#1}}
%\newcommand{\changed}[1]{}

@@

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step. 
% Author Summary not valid for PLOS ONE submissions.   
\section*{Author Summary}
  When an \changed{\emph{E. coli}} cell divides it must split into two relatively equal
  volumes in order to survive.  One mechanism that aids in this
  process is a system of proteins that oscillate between cell ends and
  help the original cell divide along its center.  A recent experiment
  found that by drastically deforming these cells this oscillation can
  be disrupted.  We examine these oscillations in similarly deformed
  cells computationally, using two widely used simulation models, and
  find good agreement using one of them.  The other model fails in a
  manner far more dramatic than has been previously observed.
  Furthermore, we find that the effect observed experimentally in
  asymmetric and flattened cells is primarily a result of the
  flattening and lateral expansion, not the asymmetry.

@@

Several models \cite{bonny2013membrane,
  halatek2012highly} modify Huang's original model to more accurately
reflect experimental findings, which were published after his 2003
paper~\cite{meacci2006mobility, loose2011min}. These models introduce
diffusion along the membrane that is two to three orders of magnitude
slower than that within the cytoplasm, and increase the rate of
cytoplasmic diffusion by a factor of about five, reflecting an
experimental measurement of these
parameters~\cite{meacci2006mobility}.  They also increase
or assume \changed{instantaneous the nucleotide exchange
rate~\cite{bonny2013membrane}}.  In addition, they revise the process
of MinD
recruitment to the membrane, removing recruitment of MinD to the
membrane by MinD-MinE complexes, behavior that has not been observed
experimentally.  The model of Bonny \emph{et al.} further more allows
MinE to remain independently attached to the membrane after
disassociating with MinD~\cite{bonny2013membrane}.
%

@@

%% biology~\cite{lawson2013spatial, robb2014stochastic,
%%   oguz2014stochastic, fu2013deterministic, rudiger2014stochastic}, and
In this paper we use Huang's original 2003
model~\cite{huang2003dynamic}, which has been widely
used~\cite{hattne2005stochastic,huang2004min,kerr2006division,
  fange2006noise} in both deterministic and stochastic variants, to
study flattened cells of $0.4\micron$ thickness that are similar to
those observed by M\"annik \emph{et al.}~\cite{mannik2009bacterial}.
This allows us to test the performance of this important model in more
extreme situations than has been done previously, furthering our
understanding of how and to what extent \changed{its predictions
  fail}.

@@

However Fig.~\ref{randst-plot-ave} leaves unclear the importance of
the flattening and accompanying lateral expansion of the cell versus
cell shape irregularity and asymmetry in the temporal regularity of
the oscillations.  Although in the movies the stadiums appear to
display, on average over long time scales, a somewhat more regular
oscillation than appears in the irregular shapes \emph{A} and
\emph{B}, the limited time range of the arrow plots (350 seconds) is
inadequate to make this comparison.  We therefore return to the
correlation function between the number of proteins in a segment at
the top of the cell and the number of proteins in a segment at the
bottom of the cell.  The ``top'' and ``bottom'' regions for computing
the correlation time were established by dividing the long axis into
thirds.  This resulted in regions at the top and bottom ranging from
19\% to 33\% of the cell volume for shapes \emph{A} and \emph{B}, with
the stadium shapes featuring 30\% of the volume in the top and bottom
regions.  Fig.~\ref{fig:corr-B} shows this plot for shape \emph{B} and
stadium \emph{B}.  Here we plot correlations of correlation times up
to 500 seconds, a biologically relevant timescale.  \changed{\emph{E. coli}} cell
doubling times vary widely depending on environmental conditions, but
will often be between 20 minutes to 100
minutes~\cite{pierucci1972chromosome}, while fluorescent microscopy
experiments have shown that the FtsZ ring builds itself with a half
life of roughly 30 seconds~\cite{stricker2002rapid}.

@@

\section*{Conclusion}
We find qualitative agreement between our stochastic simulations and
experiments showing disrupted bipolar MinD oscillation in asymmetric
flattened \changed{\emph{E. coli}} cells~\cite{mannik2012robustness}.  As observed
experimentally, our simulations predict that MinD maxima will form in
a spatially irregular sequence of locations, which follow an overall
bipolar trend.  This result builds upon existing results showing that
this model and its variants are effective in a variety of wild-type
and mutated cell shapes~\cite{fange2006noise, varma2008min,
  kruse2007experimentalist}, and reinforces that stochastic variations
of Huang's 2003 model~\cite{fange2006noise, kerr2006division} has
considerable predictive power beyond the standard wild-type cell.
